The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-12-11 Fsing/FrobeniusThresholdsPaper/FrobeniusThresholds.tex
Small changes to paper

For example, \emph{Hilbert-Kunz multiplicity} is realized in this way. 
Suppose that $I$ is an ideal of a ring $R$ of characteristic $p>0$.
Given an integer $e \geq 1$, 
$I^{[p^e]}$ denotes the $p^e$-th Frobenius power of $I$, that is, the ideal generated by the $p^e$-th powers of the elements of $I$. 
If $R$ has dimension $d$ and $\lambda\left(R/I^{[p^e]}\right)$ denotes the length of $R/I^{[p^e]}$, 
then the limit of $\lambda\left(R/I^{[p^e]}\right)/p^{ed}$ as $e$ grows is the Hilbert-Kunz multiplicity of $R$ with respect to $I$.

Consider a nonzero polynomial $f$ and a natural number $e$.  If $f$ does not vanish at the origin, then set $\nu_f^{\idealm}(p^e) := \infty$.  Otherwise, $f \in \idealm$, and we instead define
\[ \nu_f^{\idealm}(p^e) \coloneqq  \max \{ n \in \mathbb{N} : f^n \notin \idealm^{[p^e]} \}. \]
If $f^n \notin \idealm^{[p^e]}$ for some $n$, then by the flatness of Frobenius \cite{KunzCharacterizationsOfRegularLocalRings}, $f^{pn} \notin \left( \idealm^{[p^e]} \right)^{[p]} = \idealm^{[p^{e+1}]}$. 
Hence the sequence $\left\{ \frac{ \nu_f^{\idealm}(p^e)}{p^e} \right\}_{e=1}^\infty$ is non-decreasing, and since $f \in \idealm$, $f^{p^e} \in \idealm^{[p^e]}$, the sequence is bounded above by $1$.  
%where $\idealm^{[p^e]}$ denotes the $p^e$-th Frobenius power of $\idealm$, that is, the ideal generated by the $p^e$-th powers of the elements of $\idealm$.
%It is not hard to see that this number is the largest power of $f$ that is part of a free basis for the localization of the ambient polynomial ring at the origin, over its subring of $p^e$-th powers.
%As noted earlier, this number also has a natural interpretation in terms of $F$-purity of pairs.
Following our outline,  we define
\[ \ft^{\idealm}(f) \coloneqq  \lim_{e \to \infty} \frac{ \nu_f^{\idealm}(p^e)}{p^e}. \]
This limit exists by the above discussion, and is a rational number when $f\in \idealm$, though the latter is far from obvious  \cite{BlickleMustataSmithDiscretenessAndRationalityOfFThresholds}.
Inspired by its connections with the $F$-purity of pairs,  this limit is called the \emph{$F$-pure threshold} of $f$ at the origin.

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i11 : R = ZZ/5[x,y,z];

i12 : m = ideal(x, y, z);

o12 : Ideal of R

i13 : time nu(2, m, m^2) -- uses binary search (default)
     -- used 1.82479 seconds

o13 = 97

i14 : time nu(2, m, m^2, Search => Linear) -- but linear search gets luckier
     -- used 0.597035 seconds

o14 = 97
\end{MyVerbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i15 : nu(5, x^2*y^4 + y^2*z^7 + z^2*x^8, ReturnList => true)

o15 = {0, 1, 8, 44, 224, 1124}

o15 : List
\end{MyVerbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i16 : R = ZZ/11[x,y,z];

i17 : f = x^3 + y^3 + z^3 + x*y*z;

i18 : time nu(3, f) -- ContainmentTest is set to FrobeniusRoot, by default
     -- used 0.153691 seconds

o18 = 1209

i19 : time nu(3, f, ContainmentTest => StandardPower)
     -- used 10.1343 seconds

o19 = 1209
\end{MyVerbatim}
}
%\medskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i20 : R = ZZ/3[x,y];

i21 : m = ideal(x, y);

o21 : Ideal of R

i22 : nu(4, m^5)

o22 = 32

i23 : nu(4, m^5, ContainmentTest => FrobeniusPower)

o23 = 26
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i7 : R = ZZ/17[x,y,z];

i8 : f = x^3 + y^4 + z^5; -- a diagonal polynomial

i9 : time nu(10, f)
     -- used 0.0161622 seconds

o9 = 1541642394460

i10 : time nu(10, f, UseSpecialAlgorithms => false)
     -- used 2.06877 seconds

o10 = 1541642394460
\end{MyVerbatim}
}
\smallskip

>>>>

For example, the \emph{Hilbert--Kunz multiplicity} is realized in this way. 
Suppose that $I$ is an ideal of a ring $R$ of characteristic $p>0$.
Given an integer $e \geq 1$, 
$I^{[p^e]}$ denotes the $p^e$-th Frobenius power of $I$, that is, the ideal generated by the $p^e$-th powers of the elements of $I$. 
If $R$ has dimension $d$ and $\lambda\left(R/I^{[p^e]}\right)$ denotes the length of $R/I^{[p^e]}$, 
then the limit of $\lambda\left(R/I^{[p^e]}\right)/p^{ed}$ as $e$ tends to infinity is the Hilbert--Kunz multiplicity of $R$ with respect to $I$.

Consider a nonzero polynomial $f$ and a natural number $e$.  If $f$ does not vanish at the origin, then set $\nu_f^{\idealm}(p^e) \coloneqq \infty$.  Otherwise, $f \in \idealm$, and we instead define
\[ \nu_f^{\idealm}(p^e) \coloneqq  \max \{ n \in \mathbb{N} : f^n \notin \idealm^{[p^e]} \}. \]
If $f^n \notin \idealm^{[p^e]}$ for some $n$, then by the flatness of Frobenius \cite{KunzCharacterizationsOfRegularLocalRings}, $f^{pn} \notin \left( \idealm^{[p^e]} \right)^{[p]} = \idealm^{[p^{e+1}]}$. 
Hence the sequence $\Big( \frac{ \nu_f^{\idealm}(p^e)}{p^e} \Big)_{e=0}^\infty$ is non-decreasing, and since $f^{p^e} \in \idealm^{[p^e]}$, the sequence is bounded above by $1$.  
%where $\idealm^{[p^e]}$ denotes the $p^e$-th Frobenius power of $\idealm$, that is, the ideal generated by the $p^e$-th powers of the elements of $\idealm$.
%It is not hard to see that this number is the largest power of $f$ that is part of a free basis for the localization of the ambient polynomial ring at the origin, over its subring of $p^e$-th powers.
%As noted earlier, this number also has a natural interpretation in terms of $F$-purity of pairs.
Following our outline,  we define
\[ \ft^{\idealm}(f) \coloneqq  \lim_{e \to \infty} \frac{ \nu_f^{\idealm}(p^e)}{p^e}. \]
This limit exists by the above discussion, and is a rational number when $f\in \idealm$, though the latter is far from obvious  \cite[Theorem~3.1]{BlickleMustataSmithDiscretenessAndRationalityOfFThresholds}.
Inspired by its connections with the $F$-purity of pairs,  this limit is called the \emph{$F$-pure threshold} of $f$ at the origin.

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i7 : R = ZZ/5[x,y,z];

i8 : m = ideal(x, y, z);

o8 : Ideal of R

i9 : time nu(2, m, m^2) -- uses binary search (default)
     -- used 1.82479 seconds

o9 = 97

i10 : time nu(2, m, m^2, Search => Linear) -- but linear search gets luckier
     -- used 0.597035 seconds

o10 = 97
\end{MyVerbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i11 : nu(5, x^2*y^4 + y^2*z^7 + z^2*x^8, ReturnList => true)

o11 = {0, 1, 8, 44, 224, 1124}

o11 : List
\end{MyVerbatim}
}
%\medskip
\smallskip

@@

% \medskip
\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i12 : R = ZZ/11[x,y,z];

i13 : f = x^3 + y^3 + z^3 + x*y*z;

i14 : time nu(3, f) -- ContainmentTest is set to FrobeniusRoot, by default
     -- used 0.153691 seconds

o14 = 1209

i15 : time nu(3, f, ContainmentTest => StandardPower)
     -- used 10.1343 seconds

o15 = 1209
\end{MyVerbatim}
}
%\medskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i16 : R = ZZ/3[x,y];

i17 : m = ideal(x, y);

o17 : Ideal of R

i18 : nu(4, m^5)

o18 = 32

i19 : nu(4, m^5, ContainmentTest => FrobeniusPower)

o19 = 26
\end{MyVerbatim}
}
\smallskip

@@

\smallskip
{\small
\setstretch{.67}
\begin{MyVerbatim}
i20 : R = ZZ/17[x,y,z];

i21 : f = x^3 + y^4 + z^5; -- a diagonal polynomial

i22 : time nu(10, f)
     -- used 0.0161622 seconds

o22 = 1541642394460

i23 : time nu(10, f, UseSpecialAlgorithms => false)
     -- used 2.06877 seconds

o23 = 1541642394460
\end{MyVerbatim}
}
\smallskip
